---
title: GWAScat User Guide
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R/GNapi User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8](inputenc)
---

```{r knitr_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=7, fig.height=4.5,
               dev.args=list(pointsize=16))
options(width=95)
```

The [GWAScat](https://github.com/rqtl/GWAScat) package provides
access to the [NHGRI-EBI catalog of GWAS summary statistics](https://www.ebi.ac.uk/gwas).
For details on the API, see [its
documentation](https://www.ebi.ac.uk/gwas/summary-statistics/docs/),
as well as [Pjotr Prins](http://thebird.nl/)'s [documentation at
github](https://github.com/pjotrp/racket-summary-stats).


## Installation

You can install GWAScat from
[GitHub](https://github.com/rqtl/GWAScat).

You first need to install the
[devtools](https://github.com/hadley/devtools).

```{r install_devtools, eval=FALSE}
install.packages("devtools")
```

Then use `devtools::install_github()` to install GWAScat.

```{r install_with_devtools, eval=FALSE}
library(devtools)
install_github("rqtl/GWAScat")
```

Load the package with `library()`.

```{r load_package}
library(GWAScat)
```


## Lists of things

The purpose of the GWAScat package is to provide access to summary
statistics for human GWAS. First, you can get lists of studies
and traits that are available.

To get lists of studies, use `list_studies()`. The default is to
return just 20 studies. You can control that limit with the argument
`size`. You can also use `start` to step through the full set.

```{r list_studies}
list_studies(size=5)
```

To retrieve _all_ studies, set a higher limit

```{r list_all_studies}
all_studies <- list_studies(size=1000)
length(all_studies)
```

To get a list of traits, use `list_traits()`. Again the default is to
return just 20 values. To get all traits, use the `size` argument.

```{r list_all_traits}
all_traits <- list_traits(size=1000)
length(all_traits)
```

The traits are returned as identifiers like `r all_traits[1]`. To get
a description of a trait, you can use the ontology lookup service, for
example <https://www.ebi.ac.uk/efo/EFO_0001360>

Note that the traits returned are not all distinct.

```{r table_table_traits}
table(table(all_traits))
```

Chromosomes are stored as integers 1-24.

```{r list_chr}
list_chr()
```


## Get variants

To get associations for a specific variant by its rs-number, use
`get_variant()`. If you know the chromosome it is on, you'll get
faster results by providing the chromosome. And again, the default is
to return just 20 values, so use the `size` and `start` arguments if
you want a comprehensive list.

```{r get_variant}
result <- get_variant("rs2228603", 19, size=5)
result[,c("p_value", "study_accession", "trait")]
```

Use the arguments `p_lower` and `p_upper` to focus on associations
with p-value in a specified range. For example, to get all of the
associations with p-value < 10^-10^, you would do:

```{r get_variant_pupper}
result <- get_variant("rs2228603", 19, p_upper=1e-10)
result[,c("p_value", "study_accession", "trait")]
```
